---
title: "Extending the Linear Model"
output:
  xaringan::moon_reader:
    css: ["fc", "fc-fonts", "reed.css", "default"]
    lib_dir: libs
    nature:
      highlightStyle: atelier-forest-light
      highlightLines: true
      highlightSpans: true
      countIncrementalSlides: false
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.width = 8,
                      fig.align = "center",
                      message = FALSE, 
                      warning = FALSE,
                      fig.retina = 3)
library(tidyverse)
library(gganimate)
theme_set(theme_bw())
```

```{r getdata, echo = FALSE, message=FALSE}
library(DAAG)
data(allbacks)
books <- allbacks
```

# Example: shipping books
```{r plotallbacks, fig.height=5, fig.width=5}
ggplot(books, aes(x = volume, y = weight)) + 
  geom_point()
```


---
class: small

```{r fitm1, echo = FALSE}
m1 <- lm(weight ~ volume, data = books)
```

```{r allbackswline, eval=FALSE}
ggplot(books, aes(x = volume, y = weight)) + 
  geom_point() +
  geom_abline(intercept = m1$coef[1], 
              slope = m1$coef[2], col = "orchid")
```

--

```{r allbackswline-plot, ref.label="allbackswline", echo=FALSE, fig.dim=c(6, 6)}
```


---
class: small

# Fitting the linear model

```{r}
m1 <- lm(weight ~ volume, data = books)
```

--

```{r}
summary(m1)
```

---
# Multiple Regression

Allows us create a model to explain one *numerical* variable, the response, as a linear function of many explanatory variables that can be both *numerical* and
*categorical*.

--

We posit the true model:

$$Y = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \ldots + \beta_p X_p + \epsilon; \quad \epsilon \sim N(0, \sigma^2)$$

---
# Estimating $\beta_0, \beta_1$ etc.

In least-squares regression, we're still finding the estimates that minimize
the sum of squared residuals.

$$e_i = y_i - \hat{y}_i$$

$$RSS = \sum_{i = 1}^n e_i^2$$

--

And yes, they have a closed-form solution.

$$\hat{\beta} = (X'X)^{-1}X'Y$$

--

In R:

```{r eval = FALSE}
lm(Y ~ X1 + X2 + ... + Xp, data = mydata)
```


---
class: small

# Example: shipping books

```{r plotcolors, eval=FALSE}
ggplot(books, aes(x = volume,
                  y = weight, 
                  color = cover)) + 
  geom_point()
```

--

```{r plotcolors-plot, ref.label="plotcolors", echo=FALSE, fig.dim=c(7, 6)}
```


---
class: small 

# Example: shipping books

```{r}
m2 <- lm(weight ~ volume + cover, data = books)
```

--

```{r}
summary(m2)
```


---
# Example: shipping books

```{r echo = FALSE}
ggplot(books, aes(x = volume, y = weight, color = cover)) +
  geom_point() +
  geom_abline(intercept = m2$coef[1], slope = m2$coef[2], col = 2) +
  geom_abline(intercept = m2$coef[1] + m2$coef[3], slope = m2$coef[2], col = 4)
```


---
# MLR slope interpretation

The slope corresponding to the dummy variable tell us:

- How much vertical separation there is between our lines
- How much `weight` is expected to increase if `cover` goes
from 0 to 1 and `volume` is left unchanged.

Each $\hat{\beta}_i$ tells you how much you expect the $Y$ to change when you change the
$X_i$, while **holding all other variables constant**.


---
# Activity

Load in the LA homes data set and fit the following model:

$$logprice \sim logsqft + bed + city$$

```{r eval = FALSE}
URL <- "http://andrewpbray.github.io/data/LA.csv"
LA <- read.csv(URL)
```

--

1. What is the geometry of this model?

2. What appears to be the reference level for `city`?

3. In the context of this problem, what is suggested by the *sign* of the
coefficient for `bed`?  Do this make sense to you?


---
class: small

```{r}
LA <- read.csv("http://andrewpbray.github.io/data/LA.csv")
LA <- mutate(LA, logprice = log(price), logsqft = log(sqft))
m1 <- lm(logprice ~ logsqft + bed + city, data = LA)
summary(m1)
```

```{r eval = FALSE, echo = FALSE}
library(rgl)
colvec <- rep("steelblue", dim(LA)[1])
colvec[LA$city == "Long Beach"] <- "orange"
colvec[LA$city == "Santa Monica"] <- "red"
colvec[LA$city == "Westwood"] <- "green"
plot3d(x = LA$logsqft, y = LA$bed, z = LA$logprice, col = colvec, 
       xlab = "Logsqft", ylab = "# Beds", zlab = "Price")
coefs <- m1$coef
planes3d(coefs["logsqft"], coefs["bed"], -1, coefs["(Intercept)"],
         alpha = 0.4, col = "steelblue")
planes3d(coefs["logsqft"], coefs["bed"], -1, 
         coefs["(Intercept)"] + coefs["cityLong Beach"],
         alpha = 0.4, col = "orange")
planes3d(coefs["logsqft"], coefs["bed"], -1, 
         coefs["(Intercept)"] + coefs["citySanta Monica"],
         alpha = 0.4, col = "red")
planes3d(coefs["logsqft"], coefs["bed"], -1, 
         coefs["(Intercept)"] + coefs["cityWestwood"],
         alpha = 0.4, col = "green")
```

---

# Interactions

Does the relationship between `logsqft` and `logprice` change depending on 
the `city`?

.tiny[
```{r}
m2 <- lm(logprice ~ logsqft + bed + city + logsqft:city, 
         data = LA)
```
]

---
class: small

```{r}
summary(m2)
```

```{r eval = FALSE, echo = FALSE}
colvec <- rep("steelblue", dim(LA)[1])
colvec[LA$city == "Long Beach"] <- "orange"
colvec[LA$city == "Santa Monica"] <- "red"
colvec[LA$city == "Westwood"] <- "green"
plot3d(x = LA$logsqft, y = LA$bed, z = LA$logprice, col = colvec, 
       xlab = "Logsqft", ylab = "# Beds", zlab = "Price")
coefs <- m2$coef
planes3d(coefs["logsqft"], coefs["bed"], -1, coefs["(Intercept)"],
         alpha = 0.4, col = "steelblue")
planes3d(coefs["logsqft"] + coefs["logsqft:cityLong Beach"], coefs["bed"], -1, 
         coefs["(Intercept)"] + coefs["cityLong Beach"],
         alpha = 0.4, col = "orange")
planes3d(coefs["logsqft"] + coefs["logsqft:citySanta Monica"], coefs["bed"], -1, 
         coefs["(Intercept)"] + coefs["citySanta Monica"],
         alpha = 0.4, col = "red")
planes3d(coefs["logsqft"] + coefs["logsqft:cityWestwood"], coefs["bed"], -1, 
         coefs["(Intercept)"] + coefs["cityWestwood"],
         alpha = 0.4, col = "green")
```

---
# Interactions

Does the relationship between `logsqft` and `logprice` change depending on the number of `bed`?

```{r}
m3 <- lm(logprice ~ logsqft + bed + logsqft:bed, data = LA)
```

---
```{r}
summary(m3)
```

```{r echo = FALSE, eval = FALSE}
my_surface <- function(f, n=10, ...) { 
  ranges <- rgl:::.getRanges()
  x <- seq(ranges$xlim[1], ranges$xlim[2], length=n)
  y <- seq(ranges$ylim[1], ranges$ylim[2], length=n)
  z <- outer(x,y,f)
  surface3d(x, y, z, ...)
}
coefs <- coef(m3)
f <- function(x1, x2){
  coefs["(Intercept)"] + coefs["logsqft"] * x1 + coefs["bed"] * x2 + coefs["logsqft:bed"] * x1 * x2
}

x1 <- LA$logsqft
x2 <- LA$bed
plot3d(x = x1, y = x2, z = LA$logprice, col = "red", alpha = .5,
       xlab = "Logsqft", ylab = "# Beds", zlab = "Price", site = 5)
my_surface(f, alpha=.2 )
```

