---
title: "Model Diagnostics"
output:
  xaringan::moon_reader:
    css: ["fc", "fc-fonts", "reed.css", "default"]
    lib_dir: libs
    nature:
      highlightStyle: atelier-forest-light
      highlightLines: true
      highlightSpans: true
      countIncrementalSlides: false
---

```{r include = FALSE}
knitr::opts_chunk$set(fig.align = "center")
```


# A valid model

```{r}
set.seed(794)
n <- 80
x <- runif(n)
beta_1 <- 3.5
sigma <- .6
y <- rnorm(n, mean = beta_1 * x, sd = sigma)
df <- data.frame(x, y)
m1 <- lm(y ~ x, data = df)
```

---
# A valid model

```{r echo = FALSE}
library(ggplot2)
ggplot(df, aes(x = x, y = y)) + 
  geom_point(col = "steelblue") + 
  geom_abline(intercept = m1$coef[1], slope = m1$coef[2]) +
  theme_bw()
```

---
# Residual plot

```{r, fig.width = 7, fig.height = 6.5, fig.align="center"}
plot(m1, 1)
```


---
# QQ plot

```{r, fig.width = 7, fig.height = 6.5, fig.align="center"}
plot(m1, 2)
```


---
# Scale Location plot

```{r, fig.width = 7, fig.height = 6.5, fig.align="center"}
plot(m1, 3)
```


---
# Residual plot

```{r, fig.width = 7, fig.height = 6.5, fig.align="center"}
plot(m1, 4)
```


---
# Plot Quartet

```{r, fig.width = 7, fig.height = 6.5, fig.align="center"}
par(mfrow = c(2, 2))
plot(m1)
```


---
# What can break?

1. Unbiased estimation of $\hat{\beta}$
2. Accurate assessment of $SE(\hat{\beta})$
    - p-values
    - CI's
    

---
# Checking for breaks

```{r}
it <- 10000
beta_hats <- rep(NA, it)
capture <- rep(FALSE, it)
for(i in 1:it) {
  y <- rnorm(n, mean = beta_1 * x, sd = sigma)
  m <- lm(y ~ x)
  beta_hats[i] <- m$coef[2]
  ci <- confint(m)[2, ]
  capture[i] <- (ci[1] < beta_1 & beta_1 < ci[2])
}
```


---
# Checking for breaks

.tiny[
```{r fig.height = 5, warning=FALSE, message = FALSE}
ggplot(data.frame(beta_hats), aes(x = beta_hats)) +
  geom_histogram() +
  geom_vline(xintercept = mean(beta_hats), col = "tomato")
```
]

---
# Checking for breaks

### Bias in $\hat{\beta}$?

--

```{r}
mean(beta_hats) - beta_1
```

--

### Accurate assessment of $SE(\hat{\beta})$

--

```{r}
mean(capture)
```


---
# Let's break something

## Non-constant variance

```{r}
set.seed(794)
n <- 80
x <- runif(n)
y <- rnorm(n, mean = beta_1 * x, sd = .93 * x)
df <- data.frame(x, y)
m1 <- lm(y ~ x, data = df)
```

---
# Non-constant variance

```{r, echo = FALSE}
ggplot(df, aes(x = x, y = y)) + 
  geom_point(col = "steelblue") + 
  geom_abline(intercept = m1$coef[1], slope = m1$coef[2]) +
  theme_bw()
```

---
# Plot Quartet

```{r, fig.width = 7, fig.height = 7, fig.align="center", echo = FALSE}
par(mfrow = c(2, 2))
plot(m1)
```


---
# Checking for breaks

```{r}
it <- 10000
beta_hats <- rep(NA, it)
capture <- rep(FALSE, it)
for(i in 1:it) {
  y <- rnorm(n, mean = beta_1 * x, sd = .93 * x)
  m <- lm(y ~ x)
  beta_hats[i] <- m$coef[2]
  ci <- confint(m)[2, ]
  capture[i] <- (ci[1] < beta_1 & beta_1 < ci[2])
}
```


---
# Checking for breaks

.tiny[
```{r fig.height = 5, warning=FALSE, message = FALSE}
ggplot(data.frame(beta_hats), aes(x = beta_hats)) +
  geom_histogram() +
  geom_vline(xintercept = mean(beta_hats), col = "tomato")
```
]


---
# Checking for breaks

### Bias in $\hat{\beta}$?

--

```{r}
mean(beta_hats) - beta_1
```

--

### Accurate assessment of $SE(\hat{\beta})$

--

```{r}
mean(capture)
```


---
# What else can go wrong?

1. Normality of errors
2. Independence of errors
3. Linear trend in mean function
